%!TEX root = ../dissertation.tex
% the acknowledgments section

\newthought{I am grateful} to Bernhard Misof for the opportunity to
undertake this PhD project as well as for his input and support over the
five years this project took to complete.  I would not have gotten this
far without the discussions, insight, and motivation from you.

Likewise to Lars Podsiadlowski for his readiness to co-advise this
thesis and many helpful discussions.  I always appreciated your
positivity.

I also thank Dietmar Quandt as well as Torsten Wappler for their
readiness to join my commission and their time reading this thesis.

I am grateful to Christoph Mayer and Oliver Niehuis for many helpful
discussions, their careful insight, and humor.

I appreciate the financial support of the Alexander Koenig Gesellschaft,
which enabled me to attend several conferences, and the German Research
Foundation, which funded my position for the second half of my PhD
project.

Thank you to these people in no particular order for relaxed lunch
breaks, regular beach volleyball sessions, playing squash, climbing
walls, discussions or ice cream or coffee, helpful input and output,
being an outlet when I needed to vent, satisfying my gaming drive,
contributing to an exceptionally kind and welcoming atmosphere, or
simply for being a nice person:
%
Peter Grobe, Dirk Ahrens, Hans-Joachim Krammer, Matthias Geiger, Jan
Philip Oeyen, Julia Schwarzer, Sebastian Martin, Alexandros
Vasilikopoulos, Panagiotis Provataris, Jonas Astrin, Claudia Etzbauer,
Ameli Kirse, Victoria Moris, Karsten Stehr, Karoline Mauer, Juliane
Romahn, Simon K\"afer, Sofia Paraskevopoulou, Thomas Gerken, Tanja
Ziesmann, Sandra Meid, Hamideh Fard, Dirk Rohwedder, and my great
friends Jonas Eberle and Stefanie Heufelder.  I apologize to whomever I
may have forgotten here.

Gaby Nottebrock for being an awesome Master student under my
supervision.  You really made your project glow.

No thanks to Jeanne Wilbrandt and myself for wasting a year of both our
time.  It was beautiful while it lasted, though, and I learned a lot
from you.

My family has always backed me in all endeavours during my career.  I
can never convey what this unconditional support 

Thank you to my family, who has always supported me in all endeavours
during my career.  And finally, to my wife Hannah for taking care of
everything in the wrap-up stages of this work and enduring my
absent-mindedness and lack of energy for anything else.  Without you, I
would have emerged a different person.
